### 2019-03-20 _publications/jors/paper.tex
Update for stvdwtt review

%% Journal of Open Research Software Latex template -- Created By Stephen Bonner and John Brennan, Durham Universtiy, UK.

\documentclass{jors}

%% Set the header information
\pagestyle{fancy}
\definecolor{mygray}{gray}{0.6}
\renewcommand\headrule{}

\usepackage{biblatex}
\usepackage{hyperref}
\bibliography{paper}

\usepackage{graphicx}
\usepackage{hyperref}

\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%

\begin{document}

\newcommand\githublink[1]{\href{https://github.com/#1/}{\texttt{\textbf{@#1}}}}
%% {\bf Software paper for submission to the Journal of Open Research Software} \\

%% To complete this template, please replace the blue text with your own. The paper has three main sections: (1) Overview; (2) Availability; (3) Reuse potential. \\

%% Please submit the completed paper to: editor.jors@ubiquitypress.com

%% \rule{\textwidth}{1pt}

\section*{(1) Overview}

\vspace{0.5cm}

\section*{Title}
\par\bigskip
PFHub: The Phase-Field Community Hub\\[\baselineskip]
\url{https://pages.nist.gov/pfhub}

\section*{Paper Authors}

1. Wheeler, Daniel; (corresponding author) A\\
2. Keller, Trevor; A\\
3. DeWitt, Stephen J.; E\\
4. Jokisaari, Andrea M.; D\\
5. Schwen, Daniel; D\\
6. Guyer, Jonathan E.; A\\
7. Aagesen, Larry; D\\
8. Heinonen, Olle G.; B\\
9. Tonks, Michael R.; G\\
10. Voorhees, Peter W.; C\\
11. Warren, James A.; F

\section*{Paper Author Roles and Affiliations}

A. Materials Science and Engineering Division, \\
Material Measurement Laboratory, \\
National Institute of Standards and Technology,\\
Gaithersburg, MD 20899 USA

B. Argonne National Laboratory, \\
Lemont, IL 60439 USA

C. Department of Materials Science and Engineering, \\
Northwestern University, \\
Evanston, IL 60208 USA

D. Fuel Modeling and Simulation Department, \\
Idaho National Laboratory, \\
Idaho Falls, ID 83415 USA

E. Materials Science and Engineering Department, \\
University of Michigan, \\
Ann Arbor, MI 48109 USA

F. Material Measurement Laboratory Office, \\
Material Measurement Laboratory, \\
National Institute of Standards and Technology,\\
Gaithersburg, MD 20899 USA

G. Department of Materials Science and Engineering, \\
University of Florida, \\
Gainesville, FL 32611

\section*{Abstract}

An online portal provides a valuable space for scientific communities
to summarize a shared challenge, collect attempts at a solution, and
present a quantitative comparison of past attempts in a compelling
way. An exemplar of such a portal is
$\mu$MAG~\cite{mumag}~\footnote{Certain commercial equipment,
  instruments, or materials (or suppliers, or software, ...) are
  identified in this paper to foster understanding. Such
  identification does not imply recommendation or endorsement by the
  National Institute of Standards and Technology, nor does it imply
  that the materials or equipment identified are necessarily the best
  available for the purpose.\label{disclaimer}}. The reusable PFHub
framework leverages existing online services to build a static portal
website that is considerably easier to deploy and maintain without
sacrificing content or scope. The first deployment of the PFHub
framework supports phase-field practitioners and code developers
participating in an effort to improve quality assurance for
phase-field codes.

\section*{Keywords}

phase-field; materials-science; jekyll-website; reproducible-science

\section*{Introduction}

The phase-field method (PFM) describes material interfaces at the
mesoscopic scale between atomic scale models and macroscale models.
The PFM is well established and there are an assortment of code
frameworks (\emph{e.g.}, MOOSE~\cite{moose},
PRISMS-PF~\cite{prisms-pf}, FiPy~\cite{fipy},
MMSP~\cite{mmsp}~\textsuperscript{\ref{disclaimer}}) available for
solving the wide variety of phenomena associated with the PFM
(\emph{e.g.}, dendritic growth, spinodal decomposition, grain
growth). However, phase-field research groups often develop codes in
isolation and do not support or distribute the code bases to the wider
community. PFHub is a community effort spearheaded by the Center for
Hierarchical Materials Design at Northwestern University and the
National Institute of Standards and Technology to support the
development of phase-field codes. In particular, the PFHub effort
focuses on improving cross-collaboration between phase-field code
developers and practitioners by providing a standardized set of
benchmark problems~\cite{bm1, bm2} along with a web framework for
uploading and comparing benchmark results from different codes.

\section*{Implementation and architecture}

The PFHub website provides a facility for uploading, displaying and
comparing results from the benchmark problems. The website uses the
Jekyll static website generator~\cite{jekyll} along with automated
frontend processing to eliminate the need for content management
systems~\cite{cmsfree}, which are generally costly to maintain
especially for small scientific communities with limited funding and
staffing.

The workflow for uploading benchmark results relies on third party
tools using the following steps, illustrated in
Figure~\ref{fig:pfhub_website}.

\begin{itemize}
  \item The users are first required to upload simulation outputs to
    an archival resource (\emph{e.g.}, Figshare) configured with
    permissive cross-origin resource sharing (CORS).
  \item The metadata summarizing each simulation is entered into a
    form on the website, including relevant details such as memory
    usage, run time and links to the data archived in the first step.
  \item Upon submission, the Staticman app~\cite{staticman} submits
    the entered metadata as a GitHub pull request to the PFHub GitHub
    repository.  The metadata is stored in a YAML file with a unique
    path in the repository.
  \item Travis CI~\cite{travis} performs linting on the submission and
    then launches a temporary version of the proposed website using
    Surge~\cite{surge}. The PFHub admins can then examine the new
    submission and further changes can be made if necessary.
  \item Once review has been completed to the satisfaction of both the
    uploading scientist and the website maintainers, the pull request
    is merged and served to the World Wide Web using a hosting service
    compatible with GitHub Pages.
\end{itemize}

A combination of Jekyll templates and Coffeescript are used to access
and download the data links in the submitted YAML files and then
display the data in interactive plots on the website.  Previous
iterations of PFHub used the Vega JSON Library~\cite{vega} to generate
the interactive plots, but now the Plotly JavaScript Graphing
Library~\cite{plotly} is used exclusively as it provides a more
programmable interface and requires less upfront configuration.

The benchmark specifications consist of equations, narrative, plots
and code samples; and are composed in Jupyter Notbooks. The Jupyter
Notebooks are included as static objects in the website after
translation into HTML using the nbconvert tool~\cite{jupyter}.

The combination of a central repository on GitHub for website source
code and metadata with distributed data records on third-party
archives avoids the complexity and administrative overhead of
maintaining a live database and associated back-end application. The
PFHub infrastructure provides a template for other small scientific
communities to host custom content and integrate data from members of
their community.

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_website.png}
  \caption{Schematic overview of the PFHub framework for building
    scientific research portals, simply.}
  \centering
  \label{fig:pfhub_website}
\end{figure}

\section*{Quality control}

The framework has a fully automated test recipe deployed on Travis CI
with an environment built using the Nix Package
Manager~\cite{nix}. The environment is pinned to a specific version of
the Nixpkgs repository ensuring fully reproducible build and test
phases as well as ensuring that the development and automated testing
environments are identical. The full test recipe is outlined in a
\texttt{.travis.yml} file stored in the repository~\cite{travisyml}
and consists of the following steps.

\begin{itemize}
  \item Build the Nix environment from a cached storage reducing the
    build time.
  \item Run automated tests on Jupyter Notebooks using
    NBval~\cite{nbval} and Py.test~\cite{pytest}.
  \item Run validation tests on HTML files using
    HTMLProofer~\cite{htmlproofer}.
  \item Lint and test front-end Coffeescript using appropriate tools.
  \item Display a temporary version of the website using
    Surge~\cite{surge} for visual review.
\end{itemize}

\section*{(2) Availability}
\vspace{0.5cm}
\section*{Operating system}

The PFHub framework can be deployed on any platform supporting Nix,
which includes all contemporary Linux and Mac OS X platforms. Since
the framework is built with Jekyll and automated front-end processing,
it can be deployed on GitHub's Pages infrastructure, which enables
streamlined deployment without the need for any back-end
infrastructure.

\section*{Programming language}

PFHub is currently built and tested using the programming languages
and versions outlined in Table~\ref{tab:versions}.

\begin{table}[h!]
  \centering
  \caption{PFHub programming languages and corresponding supported
    versions.}
  \begin{tabular}{|l|l|}
    \hline
    Language         & Version \\
    \hline
    HTML             & 5       \\
    Jupyter Notebook & 5.4.0   \\
    JavaScript       & 5       \\
    Nix              & 2.1.3   \\
    CoffeeScript     & 1.12.7  \\
    CSS              & 4       \\
    \hline
  \end{tabular}
  \label{tab:versions}
\end{table}

\section*{Additional system requirements}

There are no additional system requirements.

\section*{Dependencies}

The entire environment can be built using the Nix Package Manager so
the only required dependency is a functional Nix installation. The
PFHub framework has over 2000 separate package dependencies using
data from the Nix package manager. The full dependency graph for PFHub
can be seen online~\cite{dependencies}.

\section*{List of contributors}

1. Wheeler, Daniel; A, \githublink{wd15} \\
2. Keller, Trevor; A, \githublink{tkphd} \\
3. DeWitt, Stephen J.; E, \githublink{stvdwtt} \\
4. Jokisaari, Andrea M.; D, \githublink{amjokisaari} \\
5. Schwen, Daniel; D, \githublink{dschwen} \\
6. Guyer, Jonathan E.; A, \githublink{guyer} \\

See the contributors list on GitHub~\cite{contributors}.

\section*{Software location:}

{\bf Archive}

\begin{description}[noitemsep,topsep=0pt]
	\item[Name:] Zenodo
	\item[Persistent identifier:]
          \href{https://dx.doi.org/10.5281/zenodo.2592705}{10.5281/zenodo.2592705}
	\item[Licence:] NIST Software License~\cite{nistlicense}
	\item[Publisher:]  Daniel Wheeler
	\item[Version published:] v0.1
	\item[Date published:] 13/03/19
\end{description}

{\bf Code repository}

\begin{description}[noitemsep,topsep=0pt]
	\item[Name:] GitHub
	\item[Persistent identifier:] \url{https://github.com/usnistgov/pfhub/tree/v0.1}
	\item[Licence:] NIST Software License~\cite{nistlicense}
	\item[Date published:] 13/03/19
\end{description}

\section*{Languages}

English

\section*{(3) Reuse potential}

The website infrastructure can be cloned as a Git repository or
downloaded as a ZIP archive and deployed with minimum effort. The
mechanism for uploading data using Staticman can be easily configured
for a new repository location. However, customizing the content of the
website for a particular scientific community would require
considerable effort. The following steps are the more challenging
aspects of deploying the framework for a new community.

\begin{itemize}
  \item Upload new benchmark definitions in a format that Jekyll can
    parse, \emph{e.g.}, Jupyter Notebook, Markdown or HTML.
  \item Edit the \texttt{benchmarks.yaml} file to reflect the new
    upload requirements and describe the figures that need to be
    generated on the upload comparison pages.
  \item Edit the \texttt{\_config.yml} file to update links and text
    related to the configuration for all aspects of the website.
  \item Update Markdown files to reflect the new content and mission
    of the scientific community.
  \item Remove data and files that are not required by the new
    community.
\end{itemize}

Currently, a deployment for a new community has not been attempted
and, thus, the above steps need to be refined and documented.

\section*{Acknowledgments}

We gratefully acknowledge input and guidance from all participants in
the series of Phase-Field workshops held between 2015 and 2018 at the
Center for Hierarchical Material Design~\cite{workshops}.

\section*{Funding statement}

D.W. wishes to acknowledge the Materials Genome Initiative funding
allocated to the National Institute of Standards and Technology. S.J.D
wishes to acknowledge funding from the U.S. Department of Energy,
Office of Basic Energy Sciences, Division of Materials Sciences and
Engineering under Award \#DE-SC0008637 as part of the Center for
PRedictive Integrated Structural Materials Science (PRISMS Center) at
University of Michigan.

\section*{Competing interests}

The authors declare that they have no competing interests.

\printbibliography

\end{document}