### 2019-08-05 _publications/jors/paper.tex
docs(publication): revise for reviewers' feedback
Implement minor revisions based on reviewer's feedback

%% Journal of Open Research Software Latex template -- Created By Stephen Bonner and John Brennan, Durham University, UK.

\documentclass{jors}

%% Set the header information
\pagestyle{fancy}
\definecolor{mygray}{gray}{0.6}
\renewcommand\headrule{}

\usepackage[sorting=none]{biblatex}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{textcomp}

\bibliography{paper}

\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%

\begin{document}

\newcommand\githublink[1]{\href{https://github.com/#1/}{\texttt{\textbf{@#1}}}}
%% {\bf Software paper for submission to the Journal of Open Research Software} \\

%% To complete this template, please replace the blue text with your own. The paper has three main sections: (1) Overview; (2) Availability; (3) Reuse potential. \\

%% Please submit the completed paper to: editor.jors@ubiquitypress.com

%% \rule{\textwidth}{1pt}

\section*{(1) Overview}

\vspace{0.5cm}

\section*{Title}
\par\bigskip
PFHub: The Phase-Field Community Hub\\[\baselineskip]

\section*{Paper Authors}

1. Wheeler, Daniel; (corresponding author) A\\
2. Keller, Trevor; A\\
3. DeWitt, Stephen J.; B\\
4. Jokisaari, Andrea M.; C\\
5. Schwen, Daniel; C\\
6. Guyer, Jonathan E.; A\\
7. Aagesen, Larry K.; C\\
8. Heinonen, Olle G.; D\\
9. Tonks, Michael R.; E\\
10. Voorhees, Peter W.; F\\
11. Warren, James A.; G

\section*{Paper Author Roles and Affiliations}

A. Materials Science and Engineering Division, \\
Material Measurement Laboratory, \\
National Institute of Standards and Technology,\\
Gaithersburg, MD 20899 USA

B. Materials Science and Engineering Department, \\
University of Michigan, \\
Ann Arbor, MI 48109 USA

C. Fuel Modeling and Simulation Department, \\
Idaho National Laboratory, \\
Idaho Falls, ID 83415 USA

D. Argonne National Laboratory, \\
Lemont, IL 60439 USA

E. Department of Materials Science and Engineering, \\
University of Florida, \\
Gainesville, FL 32611

F. Department of Materials Science and Engineering, \\
Northwestern University, \\
Evanston, IL 60208 USA

G. Material Measurement Laboratory Office, \\
Material Measurement Laboratory, \\
National Institute of Standards and Technology,\\
Gaithersburg, MD 20899 USA

\section*{Abstract}

Scientific communities struggle with the challenge of effectively and
efficiently sharing content and data. An online portal provides a
valuable space for scientific communities to discuss challenges and
collate scientific results. Examples of such portals include the
Micromagnetic Modeling Group (\textmu MAG~\cite{mumag}), the
Interatomic Potentials Repository (IPR~\cite{ipr1, ipr2}) and on a
larger scale the NIH Genetic Sequence Database
(GenBank~\cite{genbank}). In this work, we present a description of a
generic web portal that leverages existing online services to provide
a framework that may be adopted by other small scientific
communities. The first deployment of the PFHub framework supports
phase-field practitioners and code developers participating in an
effort to improve quality assurance for phase-field codes.

\section*{Keywords}

phase-field; materials-science; jekyll-website; reproducible-science

\section*{Introduction}

Generally, small scientific communities do not have the resources to
build and host dedicated web infrastructure to support their varied
content and data requirements. In particular, hosting and supporting a
complex content management system (CMS) including web servers, web
frameworks and databases requires a great deal of configuration and
long term support and funding. Furthermore, a turnkey CMS solution may
not meet requirements for most scientific communities that often use
arcane data formats and require custom data displays along with
client-side automation. The PFHub effort, instead of focusing on the
CMS tool, focuses on customizing and delivering the client-side
requirements whilst delegating back-end functionality to external
services that provide dependable APIs~\cite{cmsfree}.

The phase field method (PFM) describes material interfaces at the
mesoscopic scale between atomic scale models and macroscale
models~\cite{MOELANS2008268}. The PFM is well established and there
are an assortment of code frameworks (\emph{e.g.}, FiPy~\cite{fipy},
MMSP~\cite{mmsp}, MOOSE~\cite{moose}, PRISMS-PF~\cite{prisms-pf})
available for solving the wide variety of phenomena associated with
phase field (e.g. dendritic growth, spinodal decomposition, grain
growth)~\cite{tonks}. However, it is difficult for novice as well as
seasoned phase field practitioners to asses the capability of codes
for different phenomena without extensive prototyping and
groundwork. PFHub aims to provide a low barrier for comparing code
output data using a standard set of metrics.

PFHub is a community effort spearheaded by the Center for Hierarchical
Materials Design at Northwestern University and the National Institute
of Standards and Technology in support of phase-field code
development. The current PFHub deployment~\cite{pfhub} focuses on
improving cross-collaboration between phase-field code developers and
practitioners by providing a standardized set of benchmark
problems~\cite{bm1, bm2} and a workflow for uploading and comparing
benchmark results from different phase-field codes.

Community based scientific efforts often require web services to share
and display data in unique ways between groups and institutions. These
services are difficult to implement due to the groundwork required to
investigate and prototype the many data-sharing and CMS tools
available. The PFHub framework provides a template for other
scientific projects beyond the phase field community. The method
outlined in this paper of using static infrastructure coupled with
small independent third party web services provides a flexible
approach eliminating the initial prototyping and on-going maintenance
required for new infrastructure, while allowing developers to focus on
their unique front-end data views.

This paper presents the first deployment of the PFHub framework
including its client-side focused design, how it employs external
services and metadata about the code base. The paper describes the
relative ease with which other scientific groups might adapt the
framework for their own purposes and deploy using the fully
reproducible Nix environment~\cite{nix}.

\section*{Implementation and architecture}

The PFHub framework provides a template for other small scientific
communities to host custom content and integrate data from members of
their community. The current deployment (see
Figure~\ref{fig:pfhub_frontpage}) provides a facility for uploading,
displaying and comparing results from benchmark problems supporting
phase-field code developers and practitioners. However, the framework
and overall philosophy are broadly transferable to other communities
with some custom configuration and content generation. The framework
uses the Jekyll static website generator~\cite{jekyll} along with
automated front-end processing to eliminate the need for a
CMS~\cite{cmsfree}, which is generally costly to maintain especially
for small scientific communities with limited funding and staffing.
The framework relies on the API, WebSocket and webhook infrastructure
that underpins the modern web and allows external services to have
full-duplex communication between servers and browsers. In particular,
PFHub relies on GitHub's well maintained API and webhook functionality
for external services (such as Travis CI~\cite{travis} and
Staticman~\cite{staticman}).

The workflow for uploading benchmark results relies on third party
tools using the following steps, illustrated in
Figure~\ref{fig:pfhub_website}.
\begin{enumerate}
  \item The users are first required to upload simulation outputs to
    an archival resource (\emph{e.g.}, Figshare\footnote{Certain
      commercial equipment, instruments, or materials (or suppliers,
      or software, ...) are identified in this paper to foster
      understanding. Such identification does not imply recommendation
      or endorsement by the National Institute of Standards and
      Technology, nor does it imply that the materials or equipment
      identified are necessarily the best available for the
      purpose.\label{disclaimer}}~\cite{figshare}) configured with
    permissive cross-origin resource sharing (CORS).
  \item The metadata summarizing each simulation is entered into a
    form on the website (see Figure~\ref{fig:pfhub_upload}), including
    relevant details such as memory usage, run time and links to the
    data archived in the first step.
  \item Upon submission, the Staticman app~\cite{staticman} submits
    the entered metadata as a pull request against the PFHub
    repository hosted by GitHub. The metadata is stored in a YAML file
    with a unique path in the repository, see~\cite{upload_data} for
    an example.
  \item Travis CI~\cite{travis} performs linting on the submission and
    then launches a temporary version of the proposed website using
    Surge~\cite{surge}. The PFHub admins can then examine the new
    submission and further changes can be made if necessary.
  \item Once review has been completed to the satisfaction of both the
    uploading scientist and the website maintainers, the pull request
    is merged and served to the World Wide Web using a hosting service
    compatible with GitHub Pages.
\end{enumerate}
A combination of Jekyll templates and CoffeeScript are used to access
and download the data links in the submitted YAML files and then
display the data in interactive plots on the website. CoffeeScript is
a higher level language than JavaScript and, thus, more readable
whilst also allowing programming in a more functional manner which
makes data manipulation pipelines both more succinct and easier to
understand. The interactive plots (see Figure~\ref{fig:pfhub_results})
are displayed using the Plotly JavaScript Graphing
Library~\cite{plotly} as it provides a programmable interface and
requires minimal configuration, see~\cite{display_data} for typical
data displayed on the interactive plots.

The current deployment of PFHub has benchmark specifications
consisting of equations, narrative, plots and code samples, and are
composed in Jupyter Notebooks. The Jupyter Notebooks are included as
static objects in the website after translation into HTML using the
nbconvert tool~\cite{jupyter}. There are currently 7 benchmark
problems each with a number of variations. At the time of writing
there are 109 separate benchmark result uploads~\cite{pfhub} submitted
as pull requests and approved following code review to ensure
compatibility with the website.

The combination of a central repository on GitHub for website source
code and metadata with distributed data records on third-party
archives avoids the complexity and administrative overhead of
maintaining a live database and associated back-end application.

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_frontpage.png}
  \caption{Front page of the current version of the deployed PFHub
    website showing links to the wiki and upload pages. The hexbin
    splash provides links to phase field papers and websites. This can
    easily be modified with alternative links and images by editing a
    YAML file.}  \centering
  \label{fig:pfhub_frontpage}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_website.png}
  \caption{Schematic overview of the PFHub framework for building
    scientific research portals, simply.}
  \centering
  \label{fig:pfhub_website}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_upload.png}
  \caption{The form used to enter simulation result metadata and then
    upload to the repository via a GitHub pull request allowing the
    data to be checked before pushing to the website.}  \centering
  \label{fig:pfhub_upload}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_results.png}
  \caption{Results comparison page for benchmark 3 (Dendritic
    Growth). The data for each upload is plotted asynchronously and the
    page is not affected if a data link dies.}  \centering
  \label{fig:pfhub_results}
\end{figure}

\section*{Quality control}

The framework has a fully automated test recipe deployed on Travis CI
with an environment built using the Nix Package Manager~\cite{nix}. A
fully automated test environment using continuous integration allows
all developers and users to have common feedback on code updates and
determine the compatibility of result uploads with the deployed
website. The environment is pinned to a specific version of the Nix
Packages Collection (Nixpkgs~\cite{nixpkgs}), ensuring fully
reproducible build and test phases as well as ensuring that the
development and automated testing environments are identical. The full
test recipe is outlined in a YAML file, \texttt{.travis.yml}, file
stored in the repository~\cite{travisyml} and consists of the
following steps.

\begin{enumerate}
  \item Build the Nix environment from a persistent cache on Travis
    CI, reducing the build time.
  \item Run automated tests on Jupyter Notebooks using
    NBval~\cite{nbval} and Py.test~\cite{pytest}.
  \item Run validation tests on HTML files using
    HTMLProofer~\cite{htmlproofer}.
  \item Lint and test front-end CoffeeScript using
    Coffeelint~\cite{coffeelint} and Mocha~\cite{mocha}, respectively.
  \item Display a temporary version of the website using
    Surge~\cite{surge} for visual review.
\end{enumerate}

\section*{(2) Availability}
\vspace{0.5cm}
\section*{Operating system}

The PFHub framework can be deployed on any platform supporting Nix,
which includes all contemporary Linux and macOS platforms. Since the
framework is built with Jekyll and automated front-end processing, it
can be deployed on GitHub's Pages infrastructure, which enables
streamlined deployment without the need for any back-end
infrastructure and, thus, is largely platform independent. For
development purposes, a local installation of either Nix (on Linux or
Mac) or Docker (on Linux, Mac or Windows) is required.

\section*{Programming language}

PFHub is currently built and tested using the programming languages
and versions outlined in Table~\ref{tab:versions}.

\begin{table}[h!]
  \centering
  \caption{PFHub programming languages and corresponding supported
    versions.}
  \begin{tabular}{|l|l|}
    \hline
    Language         & Version \\
    \hline
    HTML             & 5       \\
    Jupyter Notebook & 5.4.0   \\
    JavaScript       & 5       \\
    Nix              & 2.1.3   \\
    CoffeeScript     & 1.12.7  \\
    CSS              & 4       \\
    \hline
  \end{tabular}
  \label{tab:versions}
\end{table}

\section*{Additional system requirements}

There are no additional system requirements.

\section*{Dependencies}

The entire environment can be built using the Nix Package Manager so
the only required dependency is a functional Nix installation. The
PFHub framework has over 2000 separate package dependencies using
data from the Nix package manager. The full dependency graph for PFHub
can be seen online~\cite{dependencies}.

\section*{List of contributors}

This list is for contributors to the code base, but not those that
have only uploaded output results to the website.

1. Wheeler, Daniel; A, \githublink{wd15} \\
2. Keller, Trevor; A, \githublink{tkphd} \\
3. DeWitt, Stephen J.; B, \githublink{stvdwtt} \\
4. Jokisaari, Andrea M.; C, \githublink{amjokisaari} \\
5. Schwen, Daniel; C, \githublink{dschwen} \\
6. Guyer, Jonathan E.; A, \githublink{guyer} \\

Also, see the contributors list on GitHub~\cite{contributors}.

\section*{Software location:}

{\bf Archive}

\begin{description}[noitemsep,topsep=0pt]
	\item[Name:] Zenodo
	\item[Persistent identifier:]
          \href{https://dx.doi.org/10.5281/zenodo.2592705}{10.5281/zenodo.2592705}
	\item[Licence:] NIST Software License~\cite{nistlicense}
	\item[Publisher:]  Daniel Wheeler
	\item[Version published:] v0.1
	\item[Date published:] 13/03/19
\end{description}

{\bf Code repository}

\begin{description}[noitemsep,topsep=0pt]
	\item[Name:] GitHub
	\item[Persistent identifier:] \url{https://github.com/usnistgov/pfhub/tree/v0.1}
	\item[Licence:] NIST Software License~\cite{nistlicense}
	\item[Date published:] 13/03/19
\end{description}

\section*{Languages}

English

\section*{(3) Reuse potential}

The PFHub framework can be readily adopted by other communities that
want to follow a CMS-free philosophy and use well supported external
services. The website infrastructure can be cloned as a Git repository
or downloaded as a ZIP archive and deployed with minimum effort. The
mechanism for uploading data using Staticman can be easily configured
for a new repository location. However, customizing the content of the
website for a particular scientific community would require
considerable effort. The current effort is closely integrated with
GitHub, but future deployments could be modified to use other
repository services such as GitLab or BitBucket.

The following steps are the more challenging
aspects of deploying the framework for a new community.

\begin{itemize}
  \item Upload new data upload specifications (\emph{e.g.} the
    phase-field benchmarks in the PFHub website~\cite{pfhub}) in a
    format that Jekyll can parse, \emph{e.g.}, Jupyter Notebook,
    Markdown or HTML.
  \item Edit the \texttt{benchmarks.yaml} file to reflect the new
    upload requirements and describe the figures that need to be
    generated on the upload comparison pages.
  \item Edit the \texttt{\_config.yml} file to update links and text
    related to the configuration for all aspects of the website.
  \item Update Markdown files to reflect the new content and mission
    of the scientific community.
  \item Remove data and files that are not required by the new
    community.
\end{itemize}

Further details on deployment and development of PFHub can be found in
the development guide~\cite{development_guide}.  Currently, a
deployment for a new community has not been attempted and, thus, the
above steps need to be refined and documented.

\section*{Acknowledgments}

We gratefully acknowledge input and guidance from all participants in
the series of Phase-Field workshops held between 2015 and 2018 at the
Center for Hierarchical Material Design~\cite{workshops}.

\section*{Funding statement}

D.W. wishes to acknowledge the Materials Genome Initiative funding
allocated to the National Institute of Standards and Technology. S.J.D
wishes to acknowledge funding from the U.S. Department of Energy,
Office of Basic Energy Sciences, Division of Materials Sciences and
Engineering under Award \#DE-SC0008637 as part of the Center for
PRedictive Integrated Structural Materials Science (PRISMS Center) at
University of Michigan. P.W.V. is grateful for the financial
assistance under the award 70NANB14H012 from the National Institute of
Standards and Technology as part of the Center for Hierarchical
Materials Design (CHiMaD).

\section*{Competing interests}

The authors declare that they have no competing interests.

\printbibliography

\end{document}